---
title: "Chagas_Health_Centers"
author: "Mughundhan Chandrasekar"
date: "7/07/2017"
output: html_document
---

###**INTRODUCTION**

- People often report to local "health posts", that their house is infested. We can install n (5-10) stations for the people to report.
- The dataset holds information pertaining to the houses like the geospatial coordinates, predicted probabiity of being infested etc.
- **AIM**: Build R shiny app, that will accept the number of stations as input (depends upon the budget, entered by the user) and renders a map with spatial distribution of the health posts at optimal locations.
- In this report, I have clustered the regions and idnetifed the optimal locations for installing the health facility using **K-Means Clustering technique**.

**NOTE: The data munging and feature engineering operations (along with sample visualizations) involved 500+ lines of code in R. As this project is worked for the client based out in Arequipa (Peru), the code is hidden intentionally and only a sample of the dataset is used here to make sure that the anonymity is preserved.**


###1. Creating an Environment
- Involves loading the appropriate libraries
- Load the dataset into the working environment

```{r warning=FALSE, message=FALSE}
rm(list=ls())
library(lubridate) # for csv files
library(leaflet)   # maps
library(dplyr)     # for piping purpose %>%
library(sp)
library(rgdal)
library(geosphere)
library(dismo)
library(rgeos)
library(fields)

#library(lpSolve)   # fir linear programming in R
setwd("/Users/Mughundhan/Data Science Intern/Chagas")
fdata <- read.csv("fdata.csv")
```


###2. Analyzing the Dataset
- Let us have a deeper look at the dataset in-order to gain more insights.
```{r warning=FALSE, message=FALSE, echo=FALSE}
#head(fdata)
str(fdata)
#summary(fdata)
```

###3. Data Munging

- Let us have a look at the missing values in each column
- Remove unnecessary columns
- Take a subset to work on that

```{r warning=FALSE, message=FALSE, echo=FALSE}
sapply(fdata, function(x) sum(is.na(x))) #Number of Missing Values in each column
table(is.na(fdata))
```

- We need only Geocoordinates (latitude, longitude) and the Unicode. Let us remove all other fields.

```{r warning=FALSE, message=FALSE, echo=FALSE}
drops <- c("date", "week", "LOCAL_DATETIME_new", "LOCAL_TIME","DATETIME1", "PREDICTED_COLOR","PREDICTED_PROBAB_MEAN", "PREDICTED_PROBAB","DATETIME", "TEST_DATA","STATUS_INSPECCION", "CARACT_PREDIO","DATA_ACTION", "GROUP_NAME", "USER_NAME", "UNICODE", "V1")

names(fdata)[names(fdata) == "X"] <- "id"

fdata$id <- as.character(fdata$id)
fdata <- fdata[ ,!(names(fdata) %in% drops)]
head(fdata)
```


###3. Data Visualization

- Let us plot the coordinates on google maps using leaflet and visualize their geospatial data spread.
- Interactive map: Try to zoom-in and zoom-out. Navigate within the allocated window.

```{r warning=FALSE, message=FALSE, echo=FALSE}
fdata %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(radius = 4, stroke = FALSE, fillOpacity = 0.3, color = "blue")
```


###4. Data Clustering & Visualization

- Now let us set the number of clusters or the health facilities that needs to be installed.
- Split the data points based on their geo-coordinates and assign them to each cluster.

```{r warning=FALSE, message=FALSE}
latitude<-fdata$LATITUDE
longitude<-fdata$LONGITUDE

km <- kmeans(cbind(latitude, longitude), centers = 6)
plot(longitude, latitude, col = km$cluster, pch = 20)
```

###5. Data Centers

####5.1. Identify the Health Centers

- Identify the optimal point in each clustered region to have a health facility

```{r warning=FALSE, message=FALSE, echo=FALSE}
data_centers <- km$centers
data_centers <- as.data.frame(data_centers)
id <- c("Cluster", "Cluster", "Cluster", "Cluster", "Cluster", "Cluster")
id <- as.data.frame(id)
data_centers <- cbind(id, data_centers)
#data_centers
data_centers$id <- as.character(data_centers$id)

cluster_num <- as.data.frame(km$cluster)
#nrow(cluster_num)
#head(km$cluster)
#nrow(data_centers)

leafIcons <- icons(
      iconUrl = "/Users/Mughundhan/Data Science Intern/Chagas/hospital.png", 
      iconWidth = 38, iconHeight = 38,
      iconAnchorX = 22, iconAnchorY = 38)

data_centers %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(icon=leafIcons) 
```

####5.2. Identify the Health Centers among other houses

- Assign the nearby houses to the most optimal centroid or the health facility
- Zoom-in and check the pop-up to get an idea about the type of facility, location coordinates of each facility / house.

```{r warning=FALSE, message=FALSE, echo=FALSE}

#head(data_centers)
#head(fdata)

final <- cbind(fdata, cluster_num)
head(final)
names(data_centers)[names(data_centers) == "km$cluster"] <- "Cluster_No"
names(data_centers)[names(data_centers) == "latitude"] <- "LATITUDE"
names(data_centers)[names(data_centers) == "longitude"] <- "LONGITUDE"
names(final)[names(final) == "km$cluster"] <- "Cluster_No"

data_centers$Cluster_No <- 1:6
#head(final)
#head(data_centers)
final <- rbind(final, data_centers)
final$Cluster_No <- as.character(final$Cluster_No)
#tail(final, 10)

final$Type <- "NA"
final$Type[final$id == "Cluster"] <- "Health Facility"
final$Type[final$Type == "NA"] <- "Residence / House"
tail(final, 10)
```

```{r warning=FALSE, message=FALSE, echo=FALSE}
leafIcons <- icons(
      iconUrl = ~ifelse(final$id == "Cluster", 
                        "/Users/Mughundhan/Data Science Intern/Chagas/hospital.png", 
                               "/Users/Mughundhan/Data Science Intern/Chagas/house.png"),
      
      iconWidth = 38, iconHeight = 38,
      iconAnchorX = 22, iconAnchorY = 38)

final %>%
  leaflet() %>%
  addTiles() %>%
  # addCircleMarkers(radius = 6, stroke = FALSE, fillOpacity = 0.5, popup=data_centers$id,
  #                      color = ~ifelse(final$id == "Cluster", "green", "blue"))
  addMarkers(popup = paste("Type: ", final$Type, "<br>", 
                           "Assigned to Cluster Number: ", final$Cluster_No, "<br>",
                      "Latitude: ", final$LATITUDE, "<br>",
                      "Longitude: ", final$LONGITUDE),
                       #color = ~ifelse(final$id == "Cluster", "green", "blue"),
             icon=leafIcons)      
```

- **Final Visualization: The interactive map shown below is a clustered version of the previously shown map.**
- Click the numbered bubbles repeatedly, until the cluster splits to sub groups and the logo is displayed.
- Click the logo to retreive the information pertaining to the health facility or the residence's geocoordinates.

```{r warning=FALSE, message=FALSE, echo=FALSE}
leafIcons <- icons(
      iconUrl = ~ifelse(final$id == "Cluster", 
                        "/Users/Mughundhan/Data Science Intern/Chagas/hospital.png", 
                               "/Users/Mughundhan/Data Science Intern/Chagas/house.png"),
      
      iconWidth = 38, iconHeight = 38,
      iconAnchorX = 22, iconAnchorY = 38)

final %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(clusterOptions = markerClusterOptions(),
             popup = paste("Type: ", final$Type, "<br>",
                           "Assigned to Cluster Number: ", final$Cluster_No, "<br>",
                      "Latitude: ", final$LATITUDE, "<br>",
                      "Longitude: ", final$LONGITUDE),
#             popup=final$Cluster_No,
                       #color = ~ifelse(final$id == "Cluster", "green", "blue"),
             icon=leafIcons)
```

###6. RESULT

The optimal locations for installing the health facilities are displayed:

```{r warning=FALSE, message=FALSE, echo=FALSE}
tail(final)
```